Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2015-11-17 WSSSPE3/report/WSSSPE3_review.tex
No content change.
Section 6: microscopic changes.

In WSSSPE3, we attempted to take what we learned from WSSSPE1 and WSSSPE2
in how we can collaboratively build a workshop agenda and turn that into
an ongoing community activity.  The success or failure of this effort
of this will only become apparent over time.

The workshop had two components, presentations and working groups.  The presentations,
in the first half day of the workshop, including a inspirational keynote and a set of lightning
talks.
We used lightning talks for two reasons: first, the need of some participants to have a slot
on the agenda to justify their attendance, and second, as a way to get new ideas across
to all the attendees.  We broke with the tradition of requiring the lightning talk submitter
to self-publish their papers, and instead used a common peer-review platform\footnote{\url{http://easychair.org}}, choosing to publish the speaker slides on the workshop web site
instead.

@@

\todo{this paragraph is from the wssspe2 report - to be updated...}
However, the challenge that we have discovered
since WSSSPE2 is that it is very hard to continue the breakout groups'
activities.  The WSSSPE2 participants were willing to dedicate their time to
the groups while they were at the meeting, but afterwards, they have gone
back to their (paid) jobs.  We need to determine how to tie the WSSSPE
breakout activities to people's jobs, so that they feel that continuing them
is a higher priority than it is now, perhaps through funding the participants,
or through funding coordinators for each activity, or perhaps by getting
the workshop participants to agree to a specific schedule of activities during the
workshop.

%\begin{table*}[t]
%\centering
%\caption{Top tweets tagged \#WSSSPE or \#WSSSPE3 on Sep.\ 28--29, 2015.}\label{tab:tweets}
%  \begin{scriptsize}
%  \begin{tabular}{@{}l l l l@{}}
% \toprule
%    Author  &   Tweet  & Retweets &  Favorites
%\\ \midrule
%Neil P Chue Hong  &  Getting ready for the start of the \#WSSSPE workshop & 2 & 1
%\\ & \url{http://wssspe.researchcomputing.org.uk/wssspe3/agenda/}  & &
%\\ & Video stream: \url{http://ucarconnect.ucar.edu/live?room=cg1aud} & &
%%
%\\ August Muench & Astronomy \& Software for a totally distracting Monday (2/2) => & 1 & 3
%\\ & Workshop on Sustainable SW for Science \#WSSSPE3 & & 
%\\ & \url{http://wssspe.researchcomputing.org.uk/wssspe3/agenda/} & &
%%
%\\ Neil P Chue Hong & My \#WSSSPE lightning talk on Building a Scientific Software & 6 & 3
%\\ &  Accreditation Framework: \url{http://dx.doi.org/10.6084/m9.figshare.1555925} & & 
%%
%\\ Mozilla Science Lab & At \#wssspe3 and want to learn more about our Working Open guide & 2 & 1 
%\\ &  or Contributorship badges pilot? Keep an eye out for @abbycabs! & & 
%%
%\\ Daniel S.\ Katz & \#wssspe starting! & 1 & 3
%%
%\\ Daniel S.\ Katz & @powersoffour starting \#wssspe keynote & 1 & 2
%%
%\\ Neil P Chue Hong   &  The difference between software presented at SC \& SciPy & 1 & 4
%\\  &  according to @powersoffour straw person \#wssspe & &
%%
%\\ Daniel S.\ Katz & sustainability may mean: keeps up with bug reports, adds new features, & 3 & 1
%\\ &  works on new architectures (1/2)  -- @powersoffour at \#wssspe (1/2) & & 
%%
%\\ Daniel S.\ Katz & sustainability may mean: people use it, keeps getting funded, produces & 2 & 
%\\ &  same results, transits over people (2/2)  -- @powersoffour at \#wssspe & &
%%
%\\ Nic Weber & Rec, by Matt Turk: Ash Dryden - Ethics of Unpaid Labor & 3 & 2 
%\\ & \url{http://t.co/KfMwS4VG4N} & &
%%
%\\ Nic Weber & My .02 - many more models (than 3) for sustaining sci. software. & 3 & 2
%\\ & See this pub on open-source:  & & 
%\\ & \url{http://papers.ssrn.com/sol3/papers.cfm?abstract_id=2568185} & &
%%
%\\ Erin Robinson & \#WSSSPE - Software is like a vampire. It's only dead if you really kill it. & 2 & 2
%%
%\\ Neil P Chue Hong & Absolutely great \#wssspe keynote from @powersoffour to inspire & 2 & 1 
%\\ & the workshop. Should be archived via & & 
%\\ & \url{http://ucarconnect.ucar.edu/live?room=cg1aud#.VgmpvCdIrCQ} soon & & 
%%
%\\ Daniel S.\ Katz & \#wssspe lightning talks starting - streaming at  & 5  &  1
%\\ &  \url{http://ucarconnect.ucar.edu/live?room=cg1aud} & & 
%%
%\\ Kyle Niemeyer & Participatory open source means better documentation  & 2 & 1 
%\\ & (readme, roadmap, contributing) ? @abbycabs at \#wssspe & &
%%
%\\ Erin Robinson & Interesting resource: \url{http://mozillascience.github.io/leadership-training/} & 2 & 
%\\ &  \#WSSSPE cc: @abbycabs & & 
%%
%\\ Nic Weber & @MozillaScience + @abbycabs giving look at cool resources & 2 & 3
%\\ & for leading an open source community & & 
%\\ & \url{http://mozillascience.github.io/leadership-training/} & & 
%%
%\\ Nic Weber & Some interesting sepeartion of best practices for Geosoftware dev & 2 & 
%\\ & into categories \url{https://geodynamics.org/cig/dev/best-practices} & &
%%
%\\ Neil P Chue Hong & Some statistics of software citation in geodynamics & 2 & 
%%
%\\ Daniel S.\ Katz & @jamespjh goal: make research software less rubbish; solution is RSEs & 3 & 
%%
%\\ Nic Weber & Highly relevant to \#wssspe crowd - @Impactstory team is launching & 7 & 3
%\\ & a impact of research software project \url{http://depsy.org/}  \#verycool & & 
%%
%\\ Neil P Chue Hong & @jamespjh talking about Research Software Engineer groups in the UK & 2 & 1
%\\ &  \#wssspe @ResearchSoftEng @SoftwareSaved & & 
%%
%\\ Abby Cabunoc Mayes & Wonderful time @ \#WSSSPE 3! Encouraging 2 see a diverse community & 1 & 2 
%\\ & working 2gether (\& making real plans!) \url{http://gif.co/s6v4.gif} & & 
%%
%\\ \bottomrule
%    \end{tabular}
%    \end{scriptsize}
%\end{table*}
%
%WSSSPE actively used the online social network Twitter, with hashtag
%``\#WSSSPE''. There were substantially more tweets (messages) during the days of
%the workshops WSSSPE2, WSSSPE1.1, and WSSSPE1. Out of about 670 tweets as of Apr
%18, 2015, more than 225 were about WSSSPE2 and about 180 were posted during the
%day of the workshop. Some of the main points and highlights in the meeting are
%shown in Table~\ref{tab:tweets}, which summarizes the top \#WSSSPE tweets from
%the day of workshop, selected by the metrics that number of retweets or
%favorites larger than five and the sum of two measures greater than ten.

>>>>

In WSSSPE3, we attempted to take what we learned from WSSSPE1 and WSSSPE2 in how
we can collaboratively build a workshop agenda and turn that into an ongoing
community activity. The success or failure of this effort will only
become apparent over time.

The workshop had two components, presentations and working groups. The
presentations, in the first half day of the workshop, included an inspirational
keynote and a set of lightning talks. We used lightning talks for two reasons:
first, the need of some participants to have a slot on the agenda to justify
their attendance; and second, as a way to get new ideas across to all the
attendees. We broke with the tradition of requiring the lightning talk submitters
to self-publish their papers, and instead used a common peer-review
platform\footnote{\url{http://easychair.org}}, choosing to publish their
slides on the workshop website instead.

@@

\todo{this paragraph is from the wssspe2 report - to be updated...} However, the
challenge that we have discovered since WSSSPE2 is that it is very hard to
continue the breakout groups' activities. The WSSSPE2 participants were willing
to dedicate their time to the groups while they were at the meeting, but
afterwards, they went back to their (paid) jobs. We need to determine how
to tie the WSSSPE breakout activities to people's jobs, so that they feel that
continuing them is a higher priority than it is now, perhaps through funding the
participants, or through funding coordinators for each activity, or perhaps by
getting the workshop participants to agree to a specific schedule of activities
during the workshop.
